\texttt{gwsnr}: A Python Package for Efficient SNR Calculation of Gravitational Waves

### 2024-11-05 arxiv/paper.tex
mac

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam , Anupreeta More, Harsh Narola, Ng Chung Yin (Leo), Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannuksela, Neha Singh, and David Keitel, (2024). \texttt{gwsnr}:
A Python Package for Efficient SNR Calculation of Gravitational
Waves. \textit{PendingJournal of Open Source Software}, Pending(Pending), Pending. \url{https://doi.org/10.xxxxxx/draft}}}}

@@

The \emph{\texttt{gwsnr}} Python package addresses the need for
efficient SNR computation in gravitational wave research. It
innovatively streamlines SNR calculations, enhancing accuracy and
efficiency with several advanced techniques. Firstly, it utilizes an
innovative interpolation method, employing a partial-scaling approach
for accurately interpolating the SNR of gravitational waves from
spin-less binary systems wrt to their mass parameters. Secondly, the
package features a noise-weighted inner product method, similar to that
in the \emph{\texttt{bilby}} package (Ashton, Gregory and Biscoveanu,
Sylvia and Cornish, Neil and Dal Canto, Isaac and Kumar, Prayush and
Meacher, Duncan and Middleton, Hannah and Mistry, Divyansh and Smith,
Rory and Stevenson, Tom (2022)), but enhanced with multiprocessing
capabilities. This integration allows for the parallel processing of
complex calculations, thereby expediting the SNR computation. Thirdly, a
trained Artificial Neural Network (ANN) model is incorporated for rapid
`probability of detection' (Pdet) estimation for BBH systems with spin
precession. Lastly, \emph{\texttt{gwsnr}} leverages the
\emph{\texttt{numba}}'s Just-In-Time (njit) compiler (Lam, Pitrou, and
Seibert (2022)), which optimizes performance by compiling Python code
into machine code at runtime, drastically reducing execution times.
Beyond these technical merits, \emph{\texttt{gwsnr}} stands out for its
user-friendly features and seamless integration with other related
software packages, making it not just a powerful tool but also an
accessible one for researchers. These attributes position
\emph{\texttt{gwsnr}} as an invaluable asset in gravitational wave data
analysis, particularly in simulations of detectable binary mergers,
calculation of merger rates, determining gravitational wave lensing
rates ((\textbf{ler?}), Ng et al. (2024)), and in the analysis of
selection effects within hierarchical Bayesian frameworks. The package
thus represents a significant step forward in gravitational wave
research, enabling more precise and efficient exploration of the
universe through gravitational wave observations and simulations.
Additionally, \emph{\texttt{gwsnr}} is instrumental in the
\emph{\texttt{ler}} ((\textbf{ler?})) package for calculating detectable
rates of both lensed and unlensed gravitational wave events, showcasing
its utility in advanced gravitational wave studies.

@@

>>>>

\fancyfoot[L]{\parbox[t]{0.98\headwidth}{\footnotesize{\sffamily Hemantakumar Phurailatpam, Otto Akseli Hannuksela, (2024). \texttt{gwsnr}:
A Python Package for Efficient SNR Calculation of Gravitational
Waves. \textit{PendingJournal of Open Source Software}, Pending(Pending), Pending. \url{https://doi.org/10.xxxxxx/draft}}}}

@@

The \emph{\texttt{gwsnr}} Python package addresses the need for
efficient SNR computation in gravitational wave research. It
innovatively streamlines SNR calculations, enhancing accuracy and
efficiency with several advanced techniques. Firstly, it utilizes an
innovative interpolation method, employing a partial-scaling approach
for accurately interpolating the SNR of gravitational waves from
spin-less binary systems wrt to their mass parameters. Secondly, the
package features a noise-weighted inner product method, similar to that
in the \emph{\texttt{bilby}} package (Ashton, Gregory and Biscoveanu,
Sylvia and Cornish, Neil and Dal Canto, Isaac and Kumar, Prayush and
Meacher, Duncan and Middleton, Hannah and Mistry, Divyansh and Smith,
Rory and Stevenson, Tom (2022)), but enhanced with multiprocessing
capabilities. This integration allows for the parallel processing of
complex calculations, thereby expediting the SNR computation. Thirdly, a
trained Artificial Neural Network (ANN) model is incorporated for rapid
`probability of detection' (Pdet) estimation for BBH systems with spin
precession. Lastly, \emph{\texttt{gwsnr}} leverages the
\emph{\texttt{numba}}'s Just-In-Time (njit) compiler (Lam, Pitrou, and
Seibert (2022)), which optimizes performance by compiling Python code
into machine code at runtime, drastically reducing execution times.
Beyond these technical merits, \emph{\texttt{gwsnr}} stands out for its
user-friendly features and seamless integration with other related
software packages, making it not just a powerful tool but also an
accessible one for researchers. These attributes position
\emph{\texttt{gwsnr}} as an invaluable asset in gravitational wave data
analysis, particularly in simulations of detectable binary mergers,
calculation of merger rates, determining gravitational wave lensing
rates (Phurailatpam et al. (2024), Ng et al. (2024)), and in the
analysis of selection effects within hierarchical Bayesian frameworks.
The package thus represents a significant step forward in gravitational
wave research, enabling more precise and efficient exploration of the
universe through gravitational wave observations and simulations.
Additionally, \emph{\texttt{gwsnr}} is instrumental in the
\emph{\texttt{ler}} (Phurailatpam et al. (2024)) package for calculating
detectable rates of both lensed and unlensed gravitational wave events,
showcasing its utility in advanced gravitational wave studies.

@@

\bibitem[\citeproctext]{ref-ler}
Phurailatpam, Hemantakumar, Anupreeta More, Harsh Narola, Ng Chung Yin,
Justin Janquart, Chris Van Den Broeck, Otto Akseli Hannuksela, Neha
Singh, and David Keitel. 2024. {``Ler : LVK (LIGO-Virgo-KAGRA
Collaboration) Event (Compact-Binary Mergers) Rate Calculator and
Simulator.''} \url{https://arxiv.org/abs/2407.07526}.
